Supplement to ``{\mytitle}''

### 2022-09-05 si/si.Rnw
Moved calculations to main document

Not providing a point estimate for model parameters also has the effect of making model reproduction a more difficult task.
Fortunately, the range and median values for each of the final model parameters are provided in Tables S10 and S11 of the supplement, and bivariate relationships between some of the fitted parameters are displayed in Figures S8 and S9.
We therefore use these basic summaries to reconstruct an approximation to the set of parameters that were used by the Model~1 authors to perform inference.

@@

<<sampleBetaNu, echo=FALSE, cache=TRUE>>=
set.seed(141837)
sample_beta_nu <- function(mu_beta = 5.8, mu_nu = 0.96) {
  # This function is used to sample starting values for beta and nu.
  #
  # params:
  #   mu_beta: Average of the marginal distribution for beta1. Value was fit visually.
  #   mu_nu: Average of the marginal distribution for nu. Value was fit visually.
  # returns:
  #   data.frame of size 300x2, values of beta1 and nu with truncated bivariate
  #   normal relation.

  # Determine a good value for the marginal standard deviations
  sd_beta <- (12.1-mu_beta) / qnorm(0.995)
  sd_nu <- (0.9-mu_nu) / qnorm(0.005)

  # Create covariance matrix for the bivariate-normal distribution
  cov_mat <- rbind(
    c(sd_beta^2, -0.7 * sd_beta * sd_nu),
    c(-0.7 * sd_beta * sd_nu, sd_nu^2)
  )

  # Sample from the bivariate normal distribution
  X <- MASS::mvrnorm(n = 300, mu = c(mu_beta, mu_nu), Sigma = cov_mat)
  colnames(X) <- c("beta1", "nu")

  # Identify values that fell outside of range, so they can be resampled
  bad_beta <- which(X[, 'beta1'] > 12.1 | X[, 'beta1'] < 0.04)
  bad_nu <- which(X[, 'nu'] > 1 | X[, 'nu'] < 0.9)
  n_resamp <- length(unique(c(bad_beta, bad_nu)))

  # Keep resampling values until all values fall within the range.
  while (n_resamp > 0) {
    to_resamp <- unique(c(bad_beta, bad_nu))
    X[to_resamp, ] <- MASS::mvrnorm(n = n_resamp, mu = c(mu_beta, mu_nu), Sigma = cov_mat)

    bad_beta <- which(X[, 'beta1'] > 12.1 | X[, 'beta1'] < 0.04)
    bad_nu <- which(X[, 'nu'] > 1 | X[, 'nu'] < 0.9)
    n_resamp <- length(unique(c(bad_beta, bad_nu)))
  }

  # Return results as a data.frame
  as.data.frame(X)
}

# Run the function to get values of beta1 and nu
starting_params <- sample_beta_nu()
@

<<sample_rho, echo=FALSE, cache=TRUE>>=
sample_rho <- function(beta) {
  # This function is used to get starting values for the parameter rho, based
  # on it's strong correlation with beta1 and nu. Here, we first sample beta1
  # and nu, and then sample rho based on a regression on beta1. Additionally,
  # it is noted that rho has a bi-modal distribution, with lower values of rho
  # corresponding with high probability to higher values of beta1.
  #
  # params:
  #   beta: previously sampled values of beta1 that will be used to get values
  #         of rho.
  # returns:
  #   vector of length 300 that have an appropriate relationship with the parameter
  #   beta1.

  rhos <- rep(0, 300)

  # Get slope and intercept for the regression
  slope <- (0.4/(5.1 - 6.75))
  intercpt <- -6.75 * slope

  # Determine whice mode should the value of rho belong to, based on regression
  case_when(
    beta > 6.75 ~ 0,  # Large beta --> group "0"
    # Small beta --> more likely to be group "1", with increasing probability as beta gets smaller
    beta <= 6.75 & beta > (0.95 - intercpt) / slope ~ slope * beta + intercpt,
    TRUE ~ 0.95
  ) -> probs

  # Sample the groups that rho belongs to
  rho_group1 <- rbernoulli(300, probs)
  rho_group2 <- !rho_group1

  # Keep track of how many values are in each group, used later for resampling
  n_group1 <- sum(rho_group1)
  n_group2 <- sum(rho_group2)

  # Resample values of rho that are larger than 1
  group1 <- rnorm(n_group1, mean = 1, sd = 0.05)
  n_bad1 <- sum(group1 > 1)
  while(n_bad1 > 0) {
    group1[group1 > 1] <- rnorm(n_bad1, mean = 1, sd = 0.175)
    n_bad1 <- sum(group1 > 1)
  }

  # get standard deviation for regression
  std_beta1 <- (beta[rho_group2] - mean(beta[rho_group2])) / sd(beta[rho_group2])

  group2 <- 0.25 - 0.025 * std_beta1 - (rlnorm(n_group2, sdlog = 0.2) - 1)

  # Resample values of rho in group2 as needed
  which_bad2 <- group2 < .13
  while(sum(which_bad2) > 0) {
    group2[group2 < 0.13] <- 0.25 - 0.1 * std_beta1[which_bad2] - (rlnorm(sum(which_bad2), sdlog = 0.2) - 1)

    which_bad2 <- group2 < .13
  }

  rhos[rho_group1] <- group1
  rhos[rho_group2] <- group2

  # Return the values of rho
  rhos
}

# Sample rho, using already sampled beta1.
starting_params$rho <- sample_rho(starting_params$beta1)
@

<<sample_tau, echo=FALSE, cache=TRUE>>=
sample_tau <- function(mu = 4.1) {
  # Function used to sample tau. Tau is approximately independent of the other
  # parameters.
  #
  # params:
  #   mu: mean of the truncated normal distribution
  # returns:
  #   vector of length 300, sampled values of tau from truncated normal distribution

  # Sample from normal distribution
  X <- rnorm(300, mean = mu, sd = (1.7-mu) / qnorm(0.001))

  # Get how many resamples are needed
  n_resamp <- sum(X > 5.9 | X < 1.7)

  # Keep resampling until all values fall within desired range
  while (n_resamp > 0) {
    X[X > 5.9 | X < 1.7] <- rnorm(n_resamp, mean = mu, sd = (1.7-mu) / qnorm(0.001))

    n_resamp <- sum(X > 5.9 | X < 1.7)
  }

  # Return values of tau
  X
}

# Use function to sample tau
starting_params$tau <- sample_tau()
@

<<initialValueParams, echo=FALSE, cache=TRUE>>=

# Here we note that there are additional parameters that were fit, but their
# relationship between the remaining parameters is not given in Lee et al. (2020).
# Because of this, we simply sample these parameters using a truncated normal distribution:

POP <- 10911819

E_0_probs <- dnorm(seq(1, 1752, 1), mean = 400, sd = (1752-400) / qnorm(0.99))
I_0_probs <- dnorm(seq(1, 2271, 1), mean = 500, sd = (2271-500) / qnorm(0.99))

starting_params$E_0 <- sample(1:1752, size = 300, replace = TRUE, prob = E_0_probs) / POP
starting_params$I_0 <- sample(1:2271, size = 300, replace = TRUE, prob = I_0_probs) / POP

starting_params$S_0 <- 1 - starting_params$E_0 - starting_params$I_0
@

<<RemainingParams, echo=FALSE, cache=FALSE>>=
h1_epi <- haiti1()

bounds <- tribble(
  ~param, ~lower, ~upper,
  "beta2", 1e-08, 10,
  "beta3", 1e-08, 10,
  "beta4", 1e-08, 10,
  "beta5", 1e-08, 10,
  "beta6", 1e-08, 10
)

lower <- bounds$lower
names(lower) <- bounds$param

upper <- bounds$upper
names(upper) <- bounds$param

betas <- sobol_design(
  lower = lower,
  upper = upper,
  nseq  = 300
)

epi_params <- cbind(starting_params, betas)

not_in_start_params <- names(coef(h1_epi))[!names(coef(h1_epi)) %in% colnames(epi_params)]

for (i in 1:length(not_in_start_params)) {
  epi_params[, not_in_start_params[i]] <- coef(h1_epi)[not_in_start_params[i]]
}

epi_params$sig_sq <- 0

# simply re-ordering to match model order
epi_params <- epi_params[, names(coef(h1_epi))]

epi_rw <- rw.sd(
  # Parameters close to desired values
  beta1 = 0.001,
  tau   = 0.001,
  nu    = 0.001,
  rho   = 0.001,
  # Parameters not close to desired values
  beta2 = 0.02,
  beta3 = 0.02,
  beta4 = 0.02,
  beta5 = 0.02,
  beta6 = 0.02,
  E_0   = ivp(0.1),
  I_0   = ivp(0.1)
)

epi_rw <- rw.sd(
  # Parameters close to desired values
  beta1 = 0.005,
  tau   = 0.005,
  nu    = 0.005,
  rho   = 0.005,
  # Parameters not close to desired values
  beta2 = 0.02,
  beta3 = 0.02,
  beta4 = 0.02,
  beta5 = 0.02,
  beta6 = 0.02,
  E_0   = ivp(0.1),
  I_0   = ivp(0.1)
)

registerDoParallel(20)
registerDoRNG(48963587)

# Local MIF near MLE
stew(
  file = "output/reprod_epi_MIF2.rda",
  {
    # Keep track of how long it takes
    tLocal <- system.time({
      foreach(
        i = 1:nrow(epi_params),
        .packages = c('pomp'),
        .combine = c
      ) %dopar% {

        r_params <- unlist(epi_params[i, ])
        coef(h1_epi) <- r_params
        mif2(
          h1_epi,
          Np = 100,
          Nmif = 50,
          cooling.fraction.50 = 0.5,  # Cooling set so that we will reach smaller rw.sd at the end than the end of the unit3 search.
          rw.sd = epi_rw
        ) -> m2
      } -> MIF2_search
    })
  }
)
@

<<PlotEpiDist, echo=FALSE, fig.cap="Compare to Figure S8 of Lee et al. (2020) supplement.", fig.width=7, fig.height=7, fig.align='center'>>=
epi_results <- as.data.frame(t(coef(MIF2_search)))
epi_results$logLik <- logLik(MIF2_search)

@@

<<SimulateEpi, echo=FALSE, include=FALSE, message=FALSE, cache=TRUE>>=
epi_complete <- filter(epi_results, logLik > -2200)

# First, we need to simulate the epidemic model using the fitted parameters.
registerDoParallel(20)
registerDoRNG(18599687)
foreach(
  i = 1:nrow(epi_complete),
  .combine = rbind
) %dopar% {
  guess <- unlist(epi_complete[i, ])
  sims <- simulate(h1_epi, nsim = 20, format = 'data.frame',
                   params = guess)  # 20 sims for each set of parameters
  sims$param_set <- i
  sims
} -> all_sims

@@

Following the model fitting scheme used by the Model~1 authors, we then use these resulting parameters as starting values for the endemic phase, and refit parameters using MIF2.
The resulting bivariate relations between $\rho, \tau, \beta_1$ and $\nu$ are given in Figure TODO.

<<FitEnd, echo=FALSE, message=FALSE, include=FALSE>>=
h1_end <- haiti1(period = "endemic")
start_states <- all_sims %>% filter(week == max(week))

end_rw <- rw.sd(
  beta1 = 0.05,
  tau   = 0.005,
  nu    = 0.04,
  rho   = 0.01,
  beta2 = 0.05,
  beta3 = 0.05,
  beta4 = 0.05,
  beta5 = 0.05,
  beta6 = 0.05
)

registerDoParallel(20)
registerDoRNG(21489587)

end_start <- epi_complete %>%
  select(-logLik)

# Local MIF near MLE
stew(
  file = "output/reprod_end_MIF2.rda",
  {
    foreach(
      i = 1:nrow(end_start),
      .packages = c('pomp'),
      .combine = c
    ) %dopar% {

      output <- list()
      r_params <- unlist(end_start[i, ])

      start_states %>%
        filter(param_set == i) %>%
        select(-.id, -week, -param_set) %>%
        sapply(., median, na.rm = TRUE) -> rinit_parms

      # rinit_parms["S"] <- POP - (rinit_parms["E"] + rinit_parms["I"] + rinit_parms["A"] + rinit_parms["R"])

      # Set the inital state
      new_rinit <- Csnippet(
        sprintf(
          "
          S = nearbyint(%f);
          E = nearbyint(%f);
          I = nearbyint(%f);
          A = nearbyint(%f);
          R = nearbyint(%f);
          incid = nearbyint(%f);
          foival = %f;
          Str0 = nearbyint(%f);
          Sout = nearbyint(%f);
          Sin = nearbyint(%f);
          ",
          rinit_parms['S'],
          rinit_parms['E'],
          rinit_parms['I'],
          rinit_parms['A'],
          rinit_parms['R'],
          rinit_parms['incid'],
          rinit_parms['foival'],
          rinit_parms['Str0'],
          rinit_parms['Sout'],
          rinit_parms['Sin']
        )
      )

      h1_end <- h1_end %>% pomp(
        rinit = new_rinit,
        statenames = c("S", "E", "I", "A", "R", "incid", "foival", "Str0", "Sout", "Sin")
      )

      coef(h1_end) <- r_params
      mif2(
        h1_end,
        Np = 100,
        Nmif = 100,
        cooling.type = "hyperbolic",
        cooling.fraction.50 = 0.05,  # Cooling set so that we will reach smaller rw.sd at the end than the end of the unit3 search.
        rw.sd = epi_rw
      ) -> m2

      output[['m2']] <- m2

      coef(h1_end) <- coef(m2)
      sims <- simulate(
        h1_end, nsim = 20, format = 'data.frame',
        params = coef(m2)
      )
      sims$param_set <- i

      output[['sims']] <- sims
      output
    } -> MIF2_search_end
  }
)
@

<<plotEndParams, fig.width=7, fig.height=7, fig.align='center', fig.cap="Bivariate relationships between variables after fitting endemic period. Compare to S9 of Lee et al. (2020) supplement.", cache=TRUE, echo=FALSE>>=
MIF_params <- matrix(nrow = length(MIF2_search_end) / 2, ncol = length(coef(MIF2_search_end[[1]])))
all_sims_end <- data.frame()
liks <- rep(0, nrow(end_start))
for (i in 1:nrow(end_start)) {
  MIF_params[i, ] <- coef(MIF2_search_end[[(2 * i) - 1]])
  all_sims_end <- rbind(all_sims_end, MIF2_search_end[[(2 * i)]])
  liks[i] <- logLik(MIF2_search_end[[(2 * i) - 1]])
}

colnames(MIF_params) <- names(coef(MIF2_search_end[[1]]))
end_results <- as.data.frame(MIF_params)
end_results$logLik <- liks

@@

Now that we have fit the parameters, we can simulate from the models.

@@

quants_end <- all_sims_end %>%
  filter(!param_set %in% bad_end_pars) %>%
  select(.id, week, cases) %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

>>>>

Not providing a point estimate for model parameters also has the effect of making model reproduction a more difficult task.
We therefore rely on the range and median values for each of the final model parameters are provided in Tables S10 and S11 of the supplement, and bivariate relationships between some of the fitted parameters (Figures S8 and S9) in order to replicate their results.

@@

<<Load lee Model1 Params, echo=FALSE>>=
h1_epi <- haiti1()

RL_dir <- paste0("run_level_", run_level, "/")
lee1_epi_mif <- readRDS(file = paste0("../model1/", RL_dir, "lee1_epi_fit.rds"))
@

<<Simulate Lee Model 1, echo=FALSE, cache=TRUE>>=
N_LEE_SIMS <- switch(run_level,  10,  20,   20)

registerDoRNG(18599687)
foreach(
  i = 1:length(lee1_epi_mif),
  .combine = rbind
) %dopar% {
  guess <- coef(lee1_epi_mif)
  sims <- simulate(h1_epi, nsim = N_LEE_SIMS, format = 'data.frame',
                   params = guess)  # 20 sims for each set of parameters
  sims$param_set <- i
  sims
} -> all_sims

<<PlotEpiDist, echo=FALSE, fig.cap="Compare to Figure S8 of Lee et al. (2020) supplement.", fig.width=7, fig.height=7, fig.align='center'>>=
epi_results <- as.data.frame(t(coef(lee1_epi_mif)))
epi_results$logLik <- logLik(lee1_epi_mif)

@@

<<SimulateEpi, echo=FALSE, include=FALSE, message=FALSE, cache=TRUE>>=
h1_end <- haiti1(period = "endemic")

lee1_end_mif <- readRDS(
  file = paste0("../model1/", RL_dir, "lee1_end_fit.rds")
)

@@

Following the model fitting scheme used by the Model~1 authors, we then use these resulting parameters as starting values for the endemic phase, and refit parameters using MIF2.
To do this, we use the same number of particles and MIF iterations that were used in \citet{lee20}.
The resulting bivariate relations between $\rho, \tau, \beta_1$ and $\nu$ are given in Figure~\ref{fig:plotEndParams}.

<<FitEnd, echo=FALSE, message=FALSE, include=FALSE>>=
foreach(
  i = 1:length(lee1_end_mif),
  .combine = rbind
) %dopar% {
  guess <- coef(lee1_end_mif)
  sims <- simulate(
    lee1_end_mif[[i]], nsim = N_LEE_SIMS,
    format = 'data.frame',
    params = guess
  )
  sims$param_set <- i
  sims
} -> end_sims

# quants_end <- end_sims %>%
#   select(.id, week, cases) %>%
#   group_by(week) %>%
#   summarize(
#     q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
#     q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
#     q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
#   ) %>%
#   ungroup() %>%
#   mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))
@

<<plotEndParams, fig.width=7, fig.height=7, fig.align='center', fig.cap="Bivariate relationships between variables after fitting endemic period. Compare to S9 of Lee et al. (2020) supplement.", cache=TRUE, echo=FALSE>>=

end_results <- as.data.frame(t(coef(lee1_end_mif)))
end_results$logLik <- logLik(lee1_end_mif)

@@

Now that we have fit the parameters, we can simulate from the models.
The results of these simulations are shown in Figure~\ref{fig:plotMod1Sims}.
While not a perfect replication of the results given in \citet{lee20sup}, the similarities between the model simulations, and the fact that we based our implementation of Model~1 primarily on the source code that was provided, makes us confident that we have been able to obtain a reliable representation of the model described in \citet{lee20}.

@@

quants_end <- end_sims %>%
  filter(!param_set %in% bad_end_pars) %>%
  select(.id, week, cases) %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))
